Taking Linear Logic Apart

### 2018-04-27 doc/tlla2018/main.tex
Minor fixes

The term language of \cp is a variant of the \textpi-calculus.
The variables $\tm{x}$, $\tm{y}$, and $\tm{z}$ range over channel names.
The construct $\tm{\cpLink{x}{y}}$ links two
channels~\cite{sangiorgi1996,boreale1998}, forwarding messages received on
$\tm{x}$ to $\tm{y}$ and vice versa.
The construct $\tm{\cpCut{x}{P}{Q}}$ creates a new channel $\tm{x}$, and composes
two processes, $\tm{P}$ and $\tm{Q}$, which communicate on $\tm{x}$, in parallel.
Therefore, in $\tm{\cpCut{x}{P}{Q}}$ the name $\tm{x}$ is bound in both $\tm{P}$
and $\tm{Q}$.
In $\tm{\cpRecv{x}{y}{P}}$ and $\tm{\cpSend{x}{y}{P}{Q}}$, round brackets denote
input, square brackets denote output.
We use bound output~\cite{sangiorgi1996}, meaning that both input and output
bind a new name.
In $\tm{\cpRecv{x}{y}{P}}$ the new name $\tm{y}$ is bound in $\tm{P}$.
In $\tm{\cpSend{x}{y}{P}{Q}}$, the new name $\tm{y}$ is only bound in $\tm{P}$,
while $\tm{x}$ is only bound in $\tm{Q}$.
\begin{definition}[Terms]\label{def:cp-terms}  Process terms are given by the following grammar:
  \[
    \begin{array}{rllrll}
      \tm{P}, \tm{Q}, \tm{R}
           :=& \tm{\cpLink{x}{y}}       &\text{link}
      &  \mid& \tm{\cpCut{x}{P}{Q}}     &\text{parallel composition, ``cut''}
      \\ \mid& \tm{\cpSend{x}{y}{P}{Q}} &\text{output}
      &  \mid& \tm{\cpRecv{x}{y}{P}}    &\text{input}
      \\ \mid& \tm{\cpHalt{x}}          &\text{halt}
      &  \mid& \tm{\cpWait{x}{P}}       &\text{wait}
      \\ \mid& \tm{\cpInl{x}{P}}        &\text{select left choice}
      &  \mid& \tm{\cpInr{x}{P}}        &\text{select right choice}
      \\ \mid& \tm{\cpCase{x}{P}{Q}}    &\text{offer binary choice}
      &  \mid& \tm{\cpAbsurd{x}}        &\text{offer nullary choice}
    \end{array}
  \]
\end{definition}

Terms in \cp are identified up to structural congruence, which states that
links are symmetric, and parallel compositions $\tm{\cpCut{x}{P}{Q}}$ are
associative and commutative.
\begin{definition}[Structural congruence]\label{def:cp-equiv}
  The structural congruence $\equiv$ is the congruence closure over terms which
  satisfies the following additional axioms:
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{llcll}
      \cpEquivLinkComm
      & \tm{\cpLink{x}{y}}
      & \equiv
      & \tm{\cpLink{y}{x}}
      \\
      \cpEquivCutComm
      & \tm{\cpCut{x}{P}{Q}}
      & \equiv
      & \tm{\cpCut{x}{Q}{P}}
      \\
      \cpEquivCutAss1
      & \tm{\cpCut{x}{P}{\cpCut{y}{Q}{R}}}
      & \equiv
      & \tm{\cpCut{y}{\cpCut{x}{P}{Q}}{R}}
      & \text{if }\notFreeIn{x}{R}\text{ and }\notFreeIn{y}{P}
    \end{array}
  \]
\end{definition}

@@

Reductions relate processes with their reduced forms \eg, a reduction $\reducesto{P}{Q}$ denotes that the process $\tm{P}$ can reduce to the process $\tm{Q}$ in a single step.
\begin{definition}[Reduction]\label{def:cp-reduction}
  Reductions are described by the smallest relation $\Longrightarrow$ on process terms closed under rules in \cref{fig:cp-reduction}.
  The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of 
  $\Longrightarrow$.
\end{definition}

@@

\end{definition}

@@

\end{proof}

@@

\begin{definition}[Hyper-environment]\label{def:hcp-hyper-environment}
  \(
  \ty{\mathcal{G}}, \ty{\mathcal{H}} :=
  \ty{\emptyhypercontext}
  \; \mid \;
  \ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}
  \)
\end{definition}

@@

\end{definition}

\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal{G}}$.
\end{lemma}
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}

\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal{G}}$.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}

\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\mathcal{G}}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut or a mix.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\seq[P]{\mathcal{G}}$.
  If the last rule is a cut, there are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the cut is the corresponding mix, we eliminate both as in
    \cref{thm:cp-progress};
  \item
    if the rule under the cut is an unrelated mix, we rewrite by
    \hccpEquivScopeExt{}, and recursively eliminate the cut;
  \item
    if the rule under the cut is a logical rule, we apply one of the
    $\kappa\nu$-rules.
  \end{enumerate*}
  If the last rule is mix, there are two cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the mix is a cut or another mix, we recursively eliminate it;
  \item
    if the rule under the mix is a logical rule, we apply one of the
    $\kappa{\ppar}$-rules.
  \end{enumerate*}
\end{proof}

@@

We define a translation from terms in \cp to terms in \hcp which breaks down the term constructs in \cp into their more atomic constructs in \hcp:
\begin{definition}\label{def:cp2hcp-terms}
  \[
    \begin{array}{lcllcl}
         \tm{\mtf{\cpLink{x}{y}}}
      &  := & \tm{\cpLink{x}{y}}
      &  \tm{\mtf{\cpCut{x}{P}{Q}}}
      &  := & \tm{\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}}
      \\ \tm{\mtf{\cpSend{x}{y}{P}{Q}}}
      &  := & \tm{\piBoundSend{x}{y}{(\piPar{\mtf{P}}{\mtf{Q}})}}
      &  \tm{\mtf{\cpRecv{x}{y}{P}}}
      &  := & \tm{\piRecv{x}{y}{\mtf{P}}}
      \\ \tm{\mtf{\cpHalt{x}}}
      &  := & \tm{\piBoundSend{x}{}{\piHalt}}
      &  \tm{\mtf{\cpWait{x}{P}}}
      &  := & \tm{\piRecv{x}{}{\mtf{P}}}
      \\ \tm{\mtf{\cpInl{x}{P}}}
      &  := & \tm{\cpInl{x}{\mtf{P}}}
      &  \tm{\mtf{\cpInr{x}{P}}}
      &  := & \tm{\cpInr{x}{\mtf{P}}}
      \\ \tm{\mtf{\cpCase{x}{P}{Q}}}
      &  := & \tm{\cpCase{x}{\mtf{P}}{\mtf{Q}}}
      &  \tm{\mtf{\cpAbsurd{x}}}
      &  := & \tm{\cpAbsurd{x}}
    \end{array}
  \]
\end{definition}

\begin{theorem}\label{thm:cp2hcp-typing}
  If $\seq[P]{\Gamma}$ in \cp, then $\seq[\mtf{P}]{\Gamma}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\seq[P]{\Gamma}$.
    \begin{itemize}
  \item
    Case \textsc{Cut}. We rewrite to:
    \begin{gather*}
      \begin{array}{lcl}
        \AXC{$\seq[P]{\Gamma, \tmty{x}{A}}$}
        \AXC{$\seq[Q]{\Delta, \tmty{x}{A^\bot}}$}
        \NOM{Cut}
        \BIC{$\seq[\cpCut{x}{P}{Q}]{\Gamma, \Delta}$}
        \DisplayProof
        & \Rightarrow
        & \AXC{$\seq[\mtf{P}]{\Gamma, \tmty{x}{A}}$}
          \AXC{$\seq[\mtf{Q}]{\Delta, \tmty{x}{A^\bot}}$}
          \NOM{H-Mix}
          \BIC{$\seq[\piPar{\mtf{P}}{\mtf{Q}}]{
          \Gamma, \tmty{x}{A} \hsep \Delta, \tmty{x}{A^\bot}}$}
          \NOM{H-Cut}
          \UIC{$\seq[\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}]{
          \Gamma, \Delta}$}
          \DisplayProof
      \end{array}
    \end{gather*}
  \item
    Case $(\tens)$. We rewrite as follows:
    \begin{gather*}
      \begin{array}{lcl}
        \AXC{$\seq[P]{\Gamma, \tmty{y}{A}}$}
        \AXC{$\seq[Q]{\Delta, \tmty{x}{B}}$}
        \SYM{\tens}
        \BIC{$\seq[\cpSend{x}{y}{P}{Q}]{\Gamma, \Delta, \tmty{x}{A \tens B}}$}
        \DisplayProof
        & \Rightarrow
        & \AXC{$\seq[\mtf{P}]{\Gamma, \tmty{y}{A}}$}
          \AXC{$\seq[\mtf{Q}]{\Delta, \tmty{x}{B}}$}
          \NOM{H-Mix}
          \BIC{$\seq[\piPar{\mtf{P}}{\mtf{Q}}]{
          \Gamma, \tmty{y}{A} \hsep \Delta, \tmty{x}{B}}$}
          \SYM{\tens}
          \UIC{$\seq[\piBoundSend{x}{y}{(\piPar{\mtf{P}}{\mtf{Q}})}]{
          \Gamma, \Delta, \tmty{x}{A \tens B}}$}
          \DisplayProof
      \end{array}
    \end{gather*}
  \item
    Case $(\one)$. We rewrite as follows:
    \begin{flalign*}
      \begin{array}{lcl}
        \AXC{}\SYM{\one}
        \UIC{$\seq[\cpHalt{x}]{\tmty{x}{\one}}$}
        \DisplayProof
        & \Rightarrow
        & \AXC{}
          \NOM{H-Halt}
          \UIC{$\seq[\piHalt]{\emptyhypercontext}$}
          \SYM{\one}
          \UIC{$\seq[\piBoundSend{x}{}{\piHalt}]{\tmty{x}{\one}}$}
          \DisplayProof
      \end{array}
    \end{flalign*}
  \end{itemize}
  The other cases are trivial.
\end{proof}

\begin{theorem}\label{thm:cp2hcp-equiv}
  $\tm{P}\equiv\tm{Q}$ in \cp iff $\tm{\mtf{P}}\equiv\tm{\mtf{Q}}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
% \begin{itemize}
% \item
%   Case \cpEquivCutComm.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}}
%     & \equiv & \text{by \hccpEquivMixComm}
%     \\
%     \tm{\piNew{x}{(\piPar{\mtf{Q}}{\mtf{P}})}}
%   \end{array}
%   \end{gather*}
% \item
%   Case \cpEquivCutAss1.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\mtf{P}}{\piNew{y}{(\piPar{\mtf{Q}}{\mtf{R}})}})}}
%     & \equiv & \text{%
%                by \hccpEquivNewComm, \hccpEquivMixAss1, \hccpEquivMixComm,
%                and \hccpEquivScopeExt{}}
%     \\
%     \tm{\piNew{y}{(\piPar{\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}}{\mtf{R}})}}
%   \end{array}
%   \end{gather*}
% \end{itemize}
% The other cases are trivial.
\end{proof}

\begin{theorem}\label{thm:cp2hcp-reduction}
  If $\reducesto{P}{Q}$ in \cp, then $\reducesto{\mtf{P}}{\mtf{Q}}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the the derivation of $\reducesto{P}{Q}$.
% \begin{itemize}
% \item
%   Case \cpRedBetaTensParr.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\piBoundSend{x}{y}{(\piPar{P}{Q})}}{\piRecv{x}{z}{R}})}}
%     & \Longrightarrow & \text{by \hccpRedBetaTensParr}
%     \\
%     \tm{\piNew{x}{\piNew{y}{(\piPar{(\piPar{P}{Q})}{\cpSub{y}{z}{R}})}}}
%     & \equiv & \text{%
%                by \hccpEquivNewComm, \hccpEquivMixAss1, and \hccpEquivScopeExt1}
%     \\
%     \tm{\piNew{y}{(\piPar{P}{\piNew{x}{(\piPar{Q}{\cpSub{y}{z}{R}})}})}}
%   \end{array}
%   \end{gather*}
% \item
%   Case \cpRedBetaOneBot.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\piBoundSend{x}{}{\piHalt}}{\piRecv{x}{}{\mtf{Q}}})}}
%     & \Longrightarrow & \text{by \hccpRedBetaOneBot}
%     \\
%     \tm{(\piPar{\piHalt}{\mtf{Q}})}
%     & \equiv & \text{by \hccpEquivMixHalt1}
%     \\
%     \tm{\mtf{Q}}
%   \end{array}
%   \end{gather*}
% \end{itemize}
% The other cases are trivial.
\end{proof}

\begin{theorem}\label{hcp2cp-reduction}
  If $\reducesto{\mtf{P}}{R}$ in \hcp, then there is a $\tm{Q}$ such that
  $\reducesto{P}{Q}$ in \cp and $\tm{R}\Longrightarrow^\star\tm{\mtf{Q}}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\reducesto{\mtf{P}}{R}$.
  The cases for \hccpRedAxCut1 and the $\beta$-rules are trivial.
  For the $\kappa\nu$-rules, we rewrite by the appropriate $\kappa{\ppar}$ rule,
  and vice versa.
\end{proof}

\begin{definition}\label{def:hcp-gk}
  The relation $\tm{P}\equiv_{\gamma\kappa}\tm{Q}$ is the congruence closure of
  the $\gamma$- and $\kappa$-rules for \hcp (\cref{def:hcp-reduction}).
\end{definition}

\begin{lemma}[Disentangle]\label{lem:hcp-disentangle}
  If $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ in \hcp,
  then there exist $\seq[P_1]{\Gamma_1}, \dots, \seq[P_n]{\Gamma_n}$ in \cp,
  such that
  $\tm{P}\equiv_{\gamma\kappa}\tm{(\mtf{P_1}\ppar\dots\ppar\mtf{P_n})}$.
\end{lemma}
\begin{proof}
  We repeatedly apply the $\kappa{\ppar}$-rules to move the mixes upwards.
  There are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if a mix gets stuck under a cut, it forms a \cp cut;
  \item
    if a mix gets stuck under a $(\tens)$, it forms a \cp $(\tens)$;
  \item
    otherwise, it moves all the way to the top.
  \end{enumerate*}
  Since the only way to continue after a $(\one)$ is by using a mix, all
  applications of $(\one)$ are now followed by an application of
  \textsc{H-Halt}, forming a \cp $(\one)$.
\end{proof}

We can interpret environments as types by collapsing them using a series of
pars.
\begin{definition}\label{def:bigparr}
$\ty{\bigparr(\tmty{x_1}{A_1} , \dots , \tmty{x_n}{A_n})} =  \ty{A_1 \parr \dots \parr A_n}$.
%  \begin{gather*}
%  \begin{array}{lcl}
%    \ty{\bigparr(\tmty{x_1}{A_1} , \dots , \tmty{x_n}{A_n})}
%    & = & \ty{A_1 \parr \dots \parr A_n}
%  \end{array}
%  \end{gather*}
\end{definition}

\begin{lemma}\label{lem:cp-bigparr}
  If $\seq{\Gamma}$ in \cp, then $\seq{\bigparr\Gamma}$ in \cp.
\end{lemma}
\begin{proof}
  By induction on the structure of the environment $\Gamma$.
%  \begin{itemize}
%  \item
    If $\ty{\Gamma} = \ty{A}$, then the statement is trivially true.
%  \item
    If $\ty{\Gamma} = \ty{A, B, \Gamma'}$,
    then we apply $(\parr)$ to get $\seq{A \parr B, \Gamma'}$, and apply the induction hypothesis.
%  \end{itemize}
\end{proof}

>>>>

The term language of \cp is a variant of the \textpi-calculus.
The variables $\tm{x}$, $\tm{y}$, and $\tm{z}$ range over channel names.
The construct $\tm{\cpLink{x}{y}}$ links two
channels~\cite{sangiorgi1996,boreale1998}, forwarding messages received on
$\tm{x}$ to $\tm{y}$ and vice versa.
The construct $\tm{\cpCut{x}{P}{Q}}$ creates a new channel $\tm{x}$, and composes
two processes, $\tm{P}$ and $\tm{Q}$, which communicate on $\tm{x}$, in parallel.
Therefore, in $\tm{\cpCut{x}{P}{Q}}$ the name $\tm{x}$ is bound in both $\tm{P}$
and $\tm{Q}$.
In $\tm{\cpRecv{x}{y}{P}}$ and $\tm{\cpSend{x}{y}{P}{Q}}$, round brackets denote
input, square brackets denote output.
We use bound output~\cite{sangiorgi1996}, meaning that both input and output
bind a new name.
In $\tm{\cpRecv{x}{y}{P}}$ the new name $\tm{y}$ is bound in $\tm{P}$.
In $\tm{\cpSend{x}{y}{P}{Q}}$, the new name $\tm{y}$ is only bound in $\tm{P}$,
while $\tm{x}$ is only bound in $\tm{Q}$.
\begin{definition}[Terms]\label{def:cp-terms}  Process terms are given by the following grammar:
  \[
    \begin{array}{rllrll}
      \tm{P}, \tm{Q}, \tm{R}
           :=& \tm{\cpLink{x}{y}}       &\text{link}
      &  \mid& \tm{\cpCut{x}{P}{Q}}     &\text{parallel composition, ``cut''}
      \\ \mid& \tm{\cpSend{x}{y}{P}{Q}} &\text{output}
      &  \mid& \tm{\cpRecv{x}{y}{P}}    &\text{input}
      \\ \mid& \tm{\cpHalt{x}}          &\text{halt}
      &  \mid& \tm{\cpWait{x}{P}}       &\text{wait}
      \\ \mid& \tm{\cpInl{x}{P}}        &\text{select left choice}
      &  \mid& \tm{\cpInr{x}{P}}        &\text{select right choice}
      \\ \mid& \tm{\cpCase{x}{P}{Q}}    &\text{offer binary choice}
      &  \mid& \tm{\cpAbsurd{x}}        &\text{offer nullary choice}
    \end{array}
  \]
\end{definition}\noindent

Terms in \cp are identified up to structural congruence, which states that
links are symmetric, and parallel compositions $\tm{\cpCut{x}{P}{Q}}$ are
associative and commutative.
\begin{definition}[Structural congruence]\label{def:cp-equiv}
  The structural congruence $\equiv$ is the congruence closure over terms which
  satisfies the following additional axioms:
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{llcll}
      \cpEquivLinkComm
      & \tm{\cpLink{x}{y}}
      & \equiv
      & \tm{\cpLink{y}{x}}
      \\
      \cpEquivCutComm
      & \tm{\cpCut{x}{P}{Q}}
      & \equiv
      & \tm{\cpCut{x}{Q}{P}}
      \\
      \cpEquivCutAss1
      & \tm{\cpCut{x}{P}{\cpCut{y}{Q}{R}}}
      & \equiv
      & \tm{\cpCut{y}{\cpCut{x}{P}{Q}}{R}}
      & \text{if }\notFreeIn{x}{R}\text{ and }\notFreeIn{y}{P}
    \end{array}
  \]
\end{definition}\noindent

@@

Reductions relate processes with their reduced forms \eg, a reduction $\reducesto{P}{Q}$ denotes that the process $\tm{P}$ can reduce to the process $\tm{Q}$ in a single step.
\begin{definition}[Reduction]\label{def:cp-reduction}
  Reductions are described by the smallest relation $\Longrightarrow$ on process terms closed under rules in \cref{fig:cp-reduction}.
  The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of 
  $\Longrightarrow$.
\end{definition}\noindent

@@

\end{definition}\noindent

@@

\end{proof}\noindent

@@

\begin{definition}[Hyper-environment]\label{def:hcp-hyper-environment}
  \(
  \ty{\mathcal{G}}, \ty{\mathcal{H}} :=
  \ty{\emptyhypercontext}
  \; \mid \;
  \ty{\Gamma_1 \hsep \dots \hsep \Gamma_n}
  \)
\end{definition}\noindent

@@

\end{definition}\noindent

\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal{G}}$.
\end{lemma}
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}\noindent

\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal{G}}$.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}\noindent

\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\mathcal{G}}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut or a mix.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\seq[P]{\mathcal{G}}$.
  If the last rule is a cut, there are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the cut is the corresponding mix, we eliminate both as in
    \cref{thm:cp-progress};
  \item
    if the rule under the cut is an unrelated mix, we rewrite by
    \hccpEquivScopeExt{}, and recursively eliminate the cut;
  \item
    if the rule under the cut is a logical rule, we apply one of the
    $\kappa\nu$-rules.
  \end{enumerate*}
  If the last rule is mix, there are two cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the mix is a cut or another mix, we recursively eliminate it;
  \item
    if the rule under the mix is a logical rule, we apply one of the
    $\kappa{\ppar}$-rules.
  \end{enumerate*}
\end{proof}\noindent

@@

We define a translation from terms in \cp to terms in \hcp which breaks down the term constructs in \cp into their more atomic constructs in \hcp:
\begin{definition}\label{def:cp2hcp-terms}
  \[
    \begin{array}{lcllcl}
         \tm{\mtf{\cpLink{x}{y}}}
      &  := & \tm{\cpLink{x}{y}}
      &  \tm{\mtf{\cpCut{x}{P}{Q}}}
      &  := & \tm{\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}}
      \\ \tm{\mtf{\cpSend{x}{y}{P}{Q}}}
      &  := & \tm{\piBoundSend{x}{y}{(\piPar{\mtf{P}}{\mtf{Q}})}}
      &  \tm{\mtf{\cpRecv{x}{y}{P}}}
      &  := & \tm{\piRecv{x}{y}{\mtf{P}}}
      \\ \tm{\mtf{\cpHalt{x}}}
      &  := & \tm{\piBoundSend{x}{}{\piHalt}}
      &  \tm{\mtf{\cpWait{x}{P}}}
      &  := & \tm{\piRecv{x}{}{\mtf{P}}}
      \\ \tm{\mtf{\cpInl{x}{P}}}
      &  := & \tm{\cpInl{x}{\mtf{P}}}
      &  \tm{\mtf{\cpInr{x}{P}}}
      &  := & \tm{\cpInr{x}{\mtf{P}}}
      \\ \tm{\mtf{\cpCase{x}{P}{Q}}}
      &  := & \tm{\cpCase{x}{\mtf{P}}{\mtf{Q}}}
      &  \tm{\mtf{\cpAbsurd{x}}}
      &  := & \tm{\cpAbsurd{x}}
    \end{array}
  \]
\end{definition}\noindent

\begin{theorem}\label{thm:cp2hcp-typing}
  If $\seq[P]{\Gamma}$ in \cp, then $\seq[\mtf{P}]{\Gamma}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\seq[P]{\Gamma}$.
  We show the interesting cases:
  \begin{itemize}
  \item
    Case \textsc{Cut}. We rewrite to:
    \begin{gather*}
      \begin{array}{lcl}
        \AXC{$\seq[P]{\Gamma, \tmty{x}{A}}$}
        \AXC{$\seq[Q]{\Delta, \tmty{x}{A^\bot}}$}
        \NOM{Cut}
        \BIC{$\seq[\cpCut{x}{P}{Q}]{\Gamma, \Delta}$}
        \DisplayProof
        & \Rightarrow
        & \AXC{$\seq[\mtf{P}]{\Gamma, \tmty{x}{A}}$}
          \AXC{$\seq[\mtf{Q}]{\Delta, \tmty{x}{A^\bot}}$}
          \NOM{H-Mix}
          \BIC{$\seq[\piPar{\mtf{P}}{\mtf{Q}}]{
          \Gamma, \tmty{x}{A} \hsep \Delta, \tmty{x}{A^\bot}}$}
          \NOM{H-Cut}
          \UIC{$\seq[\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}]{
          \Gamma, \Delta}$}
          \DisplayProof
      \end{array}
    \end{gather*}
  \item
    Case $(\tens)$. We rewrite as follows:
    \begin{gather*}
      \begin{array}{lcl}
        \AXC{$\seq[P]{\Gamma, \tmty{y}{A}}$}
        \AXC{$\seq[Q]{\Delta, \tmty{x}{B}}$}
        \SYM{\tens}
        \BIC{$\seq[\cpSend{x}{y}{P}{Q}]{\Gamma, \Delta, \tmty{x}{A \tens B}}$}
        \DisplayProof
        & \Rightarrow
        & \AXC{$\seq[\mtf{P}]{\Gamma, \tmty{y}{A}}$}
          \AXC{$\seq[\mtf{Q}]{\Delta, \tmty{x}{B}}$}
          \NOM{H-Mix}
          \BIC{$\seq[\piPar{\mtf{P}}{\mtf{Q}}]{
          \Gamma, \tmty{y}{A} \hsep \Delta, \tmty{x}{B}}$}
          \SYM{\tens}
          \UIC{$\seq[\piBoundSend{x}{y}{(\piPar{\mtf{P}}{\mtf{Q}})}]{
          \Gamma, \Delta, \tmty{x}{A \tens B}}$}
          \DisplayProof
      \end{array}
    \end{gather*}
  \item
    Case $(\one)$. We rewrite as follows:
    \begin{flalign*}
      \begin{array}{lcl}
        \AXC{}\SYM{\one}
        \UIC{$\seq[\cpHalt{x}]{\tmty{x}{\one}}$}
        \DisplayProof
        & \Rightarrow
        & \AXC{}
          \NOM{H-Halt}
          \UIC{$\seq[\piHalt]{\emptyhypercontext}$}
          \SYM{\one}
          \UIC{$\seq[\piBoundSend{x}{}{\piHalt}]{\tmty{x}{\one}}$}
          \DisplayProof
      \end{array}
    \end{flalign*}
  \end{itemize}
\end{proof}\noindent

\begin{theorem}\label{thm:cp2hcp-equiv}
  $\tm{P}\equiv\tm{Q}$ in \cp iff $\tm{\mtf{P}}\equiv\tm{\mtf{Q}}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
% We show the interesting cases:
% \begin{itemize}
% \item
%   Case \cpEquivCutComm.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}}
%     & \equiv & \text{by \hccpEquivMixComm}
%     \\
%     \tm{\piNew{x}{(\piPar{\mtf{Q}}{\mtf{P}})}}
%   \end{array}
%   \end{gather*}
% \item
%   Case \cpEquivCutAss1.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\mtf{P}}{\piNew{y}{(\piPar{\mtf{Q}}{\mtf{R}})}})}}
%     & \equiv & \text{%
%                by \hccpEquivNewComm, \hccpEquivMixAss1, \hccpEquivMixComm,
%                and \hccpEquivScopeExt{}}
%     \\
%     \tm{\piNew{y}{(\piPar{\piNew{x}{(\piPar{\mtf{P}}{\mtf{Q}})}}{\mtf{R}})}}
%   \end{array}
%   \end{gather*}
% \end{itemize}
\end{proof}\noindent

\begin{theorem}\label{thm:cp2hcp-reduction}
  If $\reducesto{P}{Q}$ in \cp, then $\reducesto{\mtf{P}}{\mtf{Q}}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the the derivation of $\reducesto{P}{Q}$.
% We show the interesting cases:
% \begin{itemize}
% \item
%   Case \cpRedBetaTensParr.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\piBoundSend{x}{y}{(\piPar{P}{Q})}}{\piRecv{x}{z}{R}})}}
%     & \Longrightarrow & \text{by \hccpRedBetaTensParr}
%     \\
%     \tm{\piNew{x}{\piNew{y}{(\piPar{(\piPar{P}{Q})}{\cpSub{y}{z}{R}})}}}
%     & \equiv & \text{%
%                by \hccpEquivNewComm, \hccpEquivMixAss1, and \hccpEquivScopeExt1}
%     \\
%     \tm{\piNew{y}{(\piPar{P}{\piNew{x}{(\piPar{Q}{\cpSub{y}{z}{R}})}})}}
%   \end{array}
%   \end{gather*}
% \item
%   Case \cpRedBetaOneBot.
%   \begin{gather*}
%   \begin{array}{lcl}
%     \tm{\piNew{x}{(\piPar{\piBoundSend{x}{}{\piHalt}}{\piRecv{x}{}{\mtf{Q}}})}}
%     & \Longrightarrow & \text{by \hccpRedBetaOneBot}
%     \\
%     \tm{(\piPar{\piHalt}{\mtf{Q}})}
%     & \equiv & \text{by \hccpEquivMixHalt1}
%     \\
%     \tm{\mtf{Q}}
%   \end{array}
%   \end{gather*}
% \end{itemize}
\end{proof}\noindent

\begin{theorem}\label{hcp2cp-reduction}
  If $\reducesto{\mtf{P}}{R}$ in \hcp, then there is a $\tm{Q}$ such that
  $\reducesto{P}{Q}$ in \cp and $\tm{R}\Longrightarrow^\star\tm{\mtf{Q}}$ in \hcp.
\end{theorem}
\begin{proof}
  By induction on the derivation of $\reducesto{\mtf{P}}{R}$.
  The cases for \hccpRedAxCut1 and the $\beta$-rules are trivial.
  For the $\kappa\nu$-rules, we rewrite by the appropriate $\kappa{\ppar}$ rule,
  and vice versa.
\end{proof}\noindent

\begin{definition}\label{def:hcp-gk}
  The relation $\tm{P}\equiv_{\gamma\kappa}\tm{Q}$ is the congruence closure of
  the $\gamma$- and $\kappa$-rules for \hcp.
\end{definition}\noindent

\begin{lemma}[Disentangle]\label{lem:hcp-disentangle}
  If $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ in \hcp,
  then there exist $\seq[P_1]{\Gamma_1}, \dots, \seq[P_n]{\Gamma_n}$ in \cp,
  such that
  $\tm{P}\equiv_{\gamma\kappa}\tm{(\mtf{P_1}\ppar\dots\ppar\mtf{P_n})}$.
\end{lemma}
\begin{proof}
  We repeatedly apply the $\kappa{\ppar}$-rules to move the mixes upwards.
  There are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if a mix gets stuck under a cut, it forms a \cp cut;
  \item
    if a mix gets stuck under a $(\tens)$, it forms a \cp $(\tens)$;
  \item
    otherwise, it moves all the way to the top.
  \end{enumerate*}
  Since the only way to continue after a $(\one)$ is by using a mix, all
  applications of $(\one)$ are now followed by an application of
  \textsc{H-Halt}, forming a \cp $(\one)$.
\end{proof}\noindent
We can interpret environments as types by collapsing them using a series of
pars.
\begin{definition}\label{def:bigparr}
  \(
  \ty{\bigparr(\tmty{x_1}{A_1} , \dots , \tmty{x_n}{A_n})} =
  \ty{A_1 \parr \dots \parr A_n}
  \)
\end{definition}\noindent

\begin{lemma}\label{lem:cp-bigparr}
  If $\seq{\Gamma}$ in \cp, then $\seq{\bigparr\Gamma}$ in \cp.
\end{lemma}
\begin{proof}
  By induction on the structure of the environment $\Gamma$.
  If $\ty{\Gamma} = \ty{A}$, then the statement is trivially true.
  If $\ty{\Gamma} = \ty{A, B, \Gamma'}$, then we apply $(\parr)$ to get $\seq{A
    \parr B, \Gamma'}$, and apply the induction hypothesis.
\end{proof}\noindent

We can interpret hyper-environments as types by collapsing them using a series
of tensors. In the case of the empty hyper-environment
$\ty{\emptyhypercontext}$, we interpret this as the unit of tensor.
\begin{definition}\label{def:bigtens}
  \(
  \begin{array}{lcl}
    \ty{\bigtens(\emptyhypercontext)}
    & = & \ty{\one}
    \\
    \ty{\bigtens(\Gamma_1 \hsep \dots \hsep \Gamma_n)}
    & = & \ty{\bigparr\Gamma_1 \tens \dots \tens \bigparr\Gamma_n}
  \end{array}
  \)
\end{definition}\noindent
